---
title: "Lab3"
author: "Priya Shaji"
date: "2/18/2019"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## The Normal Distribution{.tabset}


### The data

download the data

```{r}
download.file("http://www.openintro.org/stat/data/bdims.RData", destfile = "bdims.RData")

```


load the data

```{r}
load("bdims.RData")
```

```{r}
head(bdims)
mdims <- subset(bdims, sex == 1)
fdims <- subset(bdims, sex == 0)
```



The normal distribution analysis

creating density histograms
```{r}
fhgtmean <- mean(fdims$hgt)
fhgtsd   <- sd(fdims$hgt)
```



```{r}
hist(fdims$hgt, probability = TRUE,ylim = c(0, 0.06))
x <- 140:190
y <- dnorm(x = x, mean = fhgtmean, sd = fhgtsd)
lines(x = x, y = y, col = "blue")
```

```{r}
fhgtmean <- mean(fdims$hgt)
fhgtsd   <- sd(fdims$hgt)
```

Evaluating the normal distribution

```{r}
qqnorm(fdims$hgt)
qqline(fdims$hgt)
```

```{r}
sim_norm <- rnorm(n = length(fdims$hgt), mean = fhgtmean, sd = fhgtsd)
sim_norm
```



```{r}
qqnormsim(fdims$hgt)
```



Normal probabilities
```{r}
1 - pnorm(q = 182, mean = fhgtmean, sd = fhgtsd)
sum(fdims$hgt > 182) / length(fdims$hgt)
```



### Exercises{.tabset}

#### Exercise 1

Make a histogram of men’s heights and a histogram of women’s heights. How would you compare the various aspects of the two distributions?

Answer 1

Men's Height Histogram
```{r}
hist(mdims$hgt, main="Male height", xlab="cm", ylim = c(0, 80))
```


Women's Height Histogram

```{r}
hist(fdims$hgt, main="Female height", xlab="cm", ylim = c(0, 80))
```


Most of the men's height ranges between 175 cm to 180 cm with a fre quency of 80, where as most of women's height range between 165cm to 170 cm with a frequency of 75.

#### Exercise 2

Based on the this plot, does it appear that the data follow a nearly normal distribution?

Answer 2

Yes, It looks like the data follows a nearly normal distribution

#### Exercise 3

Make a normal probability plot of sim_norm. Do all of the points fall on the line? How does this plot compare to the probability plot for the real data?

Answer 3

```{r}
qqnorm(sim_norm)
qqline(sim_norm)
```

Yes, It seems like the data follows nearly normal distribution.The points are close to the line or the points are close to the real data.

#### Exercise 4

Does the normal probability plot for fdims$hgt look similar to the plots created for the simulated data? That is, do plots provide evidence that the female heights are nearly normal?

Answer 4

```{r}
qqnorm(fdims$hgt)
qqline(fdims$hgt)
```



These plots are most likely similar. There are few variations along the y axis , though not much off as to think the distributions are not normal.

#### Exercise 5

Using the same technique, determine whether or not female weights appear to come from a normal distribution.

Answer 5 

```{r}
qqnorm(fdims$wgt)
qqline(fdims$wgt)
```

The distribution for weight is right skewed . The Q-Q plot for the real data diverges from the line. Therefore the distribution does not seem normal.

#### Exercise 6

Write out two probability questions that you would like to answer; one regarding female heights and one regarding female weights. Calculate the those probabilities using both the theoretical normal distribution as well as the empirical distribution (four probabilities in all). Which variable, height or weight, had a closer agreement between the two methods?

Answer 6

Q: What is the probability that a random chosen young adult female    is shorter than 170 cm?
```{r}
pnorm(q = 170, mean = fhgtmean, sd = fhgtsd)
```

```{r}
sum(fdims$hgt < 170) / length(fdims$hgt)
```

Q: Whats the probability that a female weighs less that 50kg?
```{r}
fwgtmean <- mean(fdims$wgt)
fwgtsd   <- sd(fdims$wgt)
pnorm(q=50,mean=fwgtmean,sd=fwgtsd)
sum(fdims$wgt < 50) / length(fdims$wgt)
```
Generally, height is closer


### On Your Own{.tabset}

#### Ques_1
1: Now let’s consider some of the other variables in the body dimensions data set. Using the figures at the end of the exercises, match the histogram to its normal probability plot. All of the variables have been standardized (first subtract the mean, then divide by the standard deviation), so the units won’t be of any help. If you are uncertain based on these figures, generate the plots in R to check.

Answer 1

a. The histogram for female biiliac (pelvic) diameter (“bii.di”) belongs to normal probability plot letter: 

PLOT B

```{r}
hist(fdims$bii.di,xlab = "Female Pelvic diameter", main = "Histogram of Female Pelvic Diameter")
qqnorm(fdims$bii.di)
qqline(fdims$bii.di)

```



b. The histogram for female elbow diameter (elb.di) belongs to normal probability plot letter ____.

PLOT C

```{r}
hist(bdims$elb.di,xlab = "Female elbow diameter", main = "Histogram of Female Elbow Diameter")

qqnorm(fdims$elb.di)
qqline(fdims$elb.di)
```


c. The histogram for general age (age) belongs to normal probability plot letter ____.

PLOT D

```{r}
hist(bdims$age, xlab = "Age in years", main = "Histogram of Sample Age in Years")

qqnorm(bdims$age)
qqline(bdims$age)
```



d. The histogram for female chest depth (che.de) belongs to normal probability plot letter ____.

PLOT A

```{r}
hist(fdims$che.de,xlab = "Female chest diameter", main = "Histogram of Female Chest Diameter")
qqnorm(fdims$che.de)
qqline(fdims$che.de)
```


#### Ques_2
2. Note that normal probability plots C and D have a slight stepwise pattern.
Why do you think this is the case?

Answer 2

Likely because of the integer values provided in the data set. Age was given in integers making the jumps a bit more obvious. Perhaps for elbow diameters, many people cluster around the same diameter, since it functions in the same way for many people.


#### Ques_3

3. As you can see, normal probability plots can be used both to assess normality and visualize skewness. Make a normal probability plot for female knee diameter (kne.di). Based on this normal probability plot, is this variable left skewed, symmetric, or right skewed? Use a histogram to confirm your findings.

Answer 3

```{r}
hist(fdims$kne.di, xlab = "Female knee diameter in cm", main = "Histogram of Female Knee Diameter")
qqnorm(fdims$kne.di)
qqline(fdims$kne.di)
```

From the probability plot it appears there are less values as the quantites increase. Therefore, it appears as through it is skewed right.


