---
title: "Lab2"
output:
  word_document: default
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```


## HotHands Analysis {.tabset}
### Analyzing kobe dataset
```{r}
load("more/kobe.RData")
head(kobe)
kobe$basket[1:9]
```


Comparing data
```{r}
outcomes <- c("H", "M")
sim_basket <- sample(outcomes, size = 1, replace = TRUE)
kobe$basket
sim_basket
```


Both data sets represent the results of 133 shot attempts, each with the same 
shooting percentage of 45%. We know that our simulated data is from a shooter 
that has independent shots. That is, we know the simulated shooter does not have
a hot hand.


### Exercises Questions

1.  What does a streak length of 1 mean, i.e. how many hits and misses are in a 
    streak of 1? What about a streak length of 0?
    
Answer 1. 
A streak of length 0: No hit/ zero hit and followed by one miss.
A streak of length 1 : One hit and followed by miss.

2.  Describe the distribution of Kobe's streak lengths from the 2009 NBA finals. 
    What was his typical streak length? How long was his longest streak of baskets?
    
Answer 2. 
The distribution is left skewed 
The typical streak lenght is 0
and the longest streak is 4

```{r}
kobe_streak <- calc_streak(kobe$basket)
barplot(table(kobe_streak))
```

3.  In your simulation of flipping the unfair coin 100 times, how many flips 
    came up heads?
```{r}
sim_unfair_coin <- sample(outcomes, size = 100, replace = TRUE, prob = c(0.2, 0.8))
sim_unfair_coin
table(sim_unfair_coin)
```
    
Answer 3. 17 heads

4.  What change needs to be made to the `sample` function so that it reflects a 
    shooting percentage of 45%? Make this adjustment, then run a simulation to 
    sample 133 shots. Assign the output of this simulation to a new object called
    `sim_basket`.

```{r}
outcomes <- c("H", "M")
sim_basket <- sample(outcomes, size = 133,replace = TRUE,prob=c(0.45,0.55))
table(sim_basket)
```


### On your own

Comparing Kobe Bryant to the Independent Shooter


Using `calc_streak`, compute the streak lengths of `sim_basket`.

```{r}
sim_basket_len <- calc_streak(sim_basket)
barplot(table(sim_basket_len))
```


-   Describe the distribution of streak lengths. What is the typical streak 
    length for this simulated independent shooter with a 45% shooting percentage?
    How long is the player's longest streak of baskets in 133 shots?
    
Answer) The streak length is left skewed.
Typical streak length : 0
Longest streak of baskets in 133 shots: 5

-   If you were to run the simulation of the independent shooter a second time, 
    how would you expect its streak distribution to compare to the distribution 
    from the question above? Exactly the same? Somewhat similar? Totally 
    different? Explain your reasoning.
    
 Answer)
```{r}
sim_basket_len <- calc_streak(sim_basket)
barplot(table(sim_basket_len))
```



The distribution should be same as the probablilty between two simulations does not change and the shoots are independent of each other.


-   How does Kobe Bryant's distribution of streak lengths compare to the 
    distribution of streak lengths for the simulated shooter? Using this 
    comparison, do you have evidence that the hot hand model fits Kobe's 
    shooting patterns? Explain.
    
Answer)

Kobe Bryant's distribution of streak lengths is similar when compared to the 
distribution of streak lengths for the simulated shooter. 
Both the data has typical streak of 0.
And both data has barplots as left skewed. There is not much evidence that hot hand model fits kobe's shooting pattern. But it is analyzed by results that the shoots are independent of each other both for Kobe Bryant's distribution and simulated distribution.
