---
title: "Lab5"
author: "Priya Shaji"
date: "3/15/2019"
output:
  html_document: default
  pdf_document: default
  word_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Inference for numerical data{.tabset}

### Lab{.tabset}

#### North Carolina births

In 2004, the state of North Carolina released a large data set containing information on births recorded in this state. This data set is useful to researchers studying the relation between habits and practices of expectant mothers and the birth of their children. We will work with a random sample of observations from this data set.

#### Exploratory analysis

Load the nc data set into our workspace.

1) Download the file and Load the data


```{r}
library(DATA606)
download.file("http://www.openintro.org/stat/data/nc.RData", destfile = "nc.RData")
load("/Users/priyashaji/Documents/cuny msds/Spring'19/data 606/Labs/Lab5/nc.RData")

```


We have observations on 13 different variables, some categorical and some numerical. The meaning of each variable is as follows.

variable	description

1) `fage`	 father’s age in years.

2) `mage`	mother’s age in years.

3) `mature`	maturity status of mother.

4) `weeks`	length of pregnancy in weeks.

5) `premie`	whether the birth was classified as premature (premie) or full-term.

6) `visits`	number of hospital visits during pregnancy.

7) `marital`	whether mother is married or not married at birth.

8) `gained`	weight gained by mother during pregnancy in pounds.

9) `weight`	weight of the baby at birth in pounds.

10) `lowbirthweight`	whether baby was classified as low birthweight (low) or not (not low).

11) `gender`	gender of the baby, female or male.

12) `habit`	status of the mother as a nonsmoker or a smoker.

13) `whitemom`	whether mom is white or not white.


As a first step in the analysis, we should consider summaries of the data. This can be done using the summary command:


```{r}
summary(nc)
```


As you review the variable summaries, consider which variables are categorical and which are numerical. For numerical variables, are there outliers? If you aren’t sure or want to take a closer look at the data, make a graph.

Consider the possible relationship between a mother’s smoking habit and the weight of her baby. Plotting the data is a useful first step because it helps us quickly visualize trends, identify strong associations, and develop research questions.


Answer) 
`fage`	 father’s age in years. : `Numerical`
`mage`	mother’s age in years. : `Numerical`
`mature`	maturity status of mother. : `Categorical`
`weeks`	length of pregnancy in weeks. : `Numerical`
`premie`	whether the birth was classified as premature (premie) or full-term. : `Categorical`
`visits`	number of hospital visits during pregnancy. :`Numerical`
`marital`	whether mother is married or not married at birth. :`Categorical`
`gained`	weight gained by mother during pregnancy in pounds. :`Numerical`
`weight`	weight of the baby at birth in pounds. : `Numerical`
`lowbirthweight`	whether baby was classified as low birthweight (low) or not (not low). : `Categorical`
`gender`	gender of the baby, female or male. : `Categorical`
`habit`	status of the mother as a nonsmoker or a smoker.: `Categorical`
`whitemom`	whether mom is white or not white. : `Categorical`

Plotting a graph between a mother’s smoking habit and the weight of her baby :

```{r}
plot(nc$habit,nc$weight)
```


non-smoker moms has few outliers for babies weighing 11 pounds and more which tells us their is less or weak association between a non smoking mother and high weight of her baby whihc is a very less likely event. According to the graph we can see, there is a strong association between mom's who are non-smoker and the weight of their babies who weigh between approx 0.5 to 4 pounds.

Research question: 

Is the baby of a mother who smokes has lower chances of being born healthy?
Does the baby of a motther who smokes has higher chances of risk of diseases and incresed risk of disability?


#### Inference

Next, we introduce a new function, inference, that we will use for conducting hypothesis tests and constructing confidence intervals.

```{r}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```


Let's pause for a moment to go through the arguments of this custom function. 
The first argument is `y`, which is the response variable that we are 
interested in: `nc$weight`. The second argument is the explanatory variable, 
`x`, which is the variable that splits the data into two groups, smokers and 
non-smokers: `nc$habit`. The third argument, `est`, is the parameter we're 
interested in: `"mean"` (other options are `"median"`, or `"proportion"`.) Next
we decide on the `type` of inference we want: a hypothesis test (`"ht"`) or a 
confidence interval (`"ci"`). When performing a hypothesis test, we also need 
to supply the `null` value, which in this case is `0`, since the null 
hypothesis sets the two population means equal to each other. The `alternative`
hypothesis can be `"less"`, `"greater"`, or `"twosided"`. Lastly, the `method` 
of inference can be `"theoretical"` or `"simulation"` based.



### Exercises{.tabset}

#### Exercise 1

What are the cases in this data set? How many cases are there in our sample?

Answer 1)

```{r}
nrow(nc)
```

Generally, In statiscs, cases are called as observations of the dataset.
Therefore, in nc dataset there are `1000` rows in a dataset

#### Exercise 2

Make a side-by-side boxplot of habit and weight. What does the plot highlight about the relationship between these two variables?

Answer 2)

```{r}
plot(nc$habit,nc$weight)
```


non-smoker moms has few outliers for babies weighing 11 pounds and more which tells us their is less or weak association between a non smoking mother and high weight of her baby whihc is a very less likely event. According to the graph we can see, there is a strong association between mom's who are non-smoker and the weight of their babies who weigh between approx 0.5 to 4 pounds.


The box plots show how the medians of the two distributions compare, but we can also compare the means of the distributions using the following function to split the weight variable into the habit groups, then take the mean of each using the mean function.

```{r}
by(nc$weight, nc$habit, mean)
```


#### Exercise 3


Check if the conditions necessary for inference are satisfied. Note that you will need to obtain sample sizes to check the conditions. You can compute the group size using the same by command above but replacing mean with length.


Answer 3) 


```{r}
by(nc$weight, nc$habit, length)

```


Let's check if the distribution is normal

```{r}
g <- ggplot(nc, aes(x = weight, colour = habit)) + geom_density()
g
```

From the above we can note the sample size is sufficiently large (>30), and their distributions approach normal, although left-skewed. We can say that the events are independent.


#### Exercise 4


Write the hypotheses for testing if the average weights of babies born to smoking and non-smoking mothers are different.

Answer 4)

```{r inf-weight-habit-ht, eval=TRUE, tidy=FALSE}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```



#### Exercise 5

Change the type argument to "ci" to construct and record a confidence interval for the difference between the weights of babies born to smoking and non-smoking mothers.

Answer 5)

By default the function reports an interval for (μnonsmoker−μsmoker) . We can easily change this order by using the order argument:

```{r}
inference(y = nc$weight, x = nc$habit, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical", 
          order = c("smoker","nonsmoker"))
```



### On Your Own{.tabset}

#### Question 1

Calculate a 95% confidence interval for the average length of pregnancies (weeks) and interpret it in context. Note that since you’re doing inference on a single population parameter, there is no explanatory variable, so you can omit the x variable from the function.


Answer 1)


```{r}
inference(y = nc$weeks, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical", 
          order = c("smoker","nonsmoker"))
```


#### Question 2

Calculate a new confidence interval for the same parameter at the 90% 
confidence level. You can change the confidence level by adding a new argument 
to the function: `conflevel = 0.90`.

Answer 2)

```{r}
inference(y = nc$weeks, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical", 
          order = c("smoker","nonsmoker"),conflevel = 0.90)
```


#### Question 3

Conduct a hypothesis test evaluating whether the average weight gained by 
younger mothers is different than the average weight gained by mature mothers.

Answer 3)

```{r}
inference(y = nc$gained, x = nc$mature, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```


#### Question 4

Now, a non-inference task: Determine the age cutoff for younger and mature 
mothers. Use a method of your choice, and explain how your method works.

Answer 4)

```{r}
by(nc$mage, nc$mature, summary)
```


```{r}
boxplot(nc$mage[nc$mature == "mature mom"])
```


I have used the method by() and boxplot() method to get the age cutoff for younger and mature 
mothers. nc$mature: mature mom , therefore for mature mom's the minimum age cutoff is 35 and for younger mom the minimum age cutoff is 13.

using the by() method:

compare the summary of the distributions using the following function to split the mage variable into the maturity groups, then take the summarize each using the summary function.

using the boxplot() method:

by plotting the boxplot for the mature mom, we get the minimum age as 35

```{r}
boxplot(nc$mage[nc$mature == "younger mom"])
```


by plotting the boxplot for the younger mom, we get the minimum age as 13



#### Question 5

Pick a pair of numerical and categorical variables and come up with a 
research question evaluating the relationship between these variables. 
Formulate the question in a way that it can be answered using a hypothesis test
and/or a confidence interval. Answer your question using the `inference` 
function, report the statistical results, and also provide an explanation in 
plain language.


Answer 5)

numerical variable:

`weeks`          | length of pregnancy in weeks.

categorical variable:

`habit`          | status of the mother as a `nonsmoker` or a `smoker`.


Research Question:

Conduct a hypothesis test evaluating how length of pregnency in weeks depend on the status of the mother as a smoker or nonsmoker.

```{r}
inference(y = nc$weeks, x = nc$habit, est = "mean", type = "ht", null = 0, 
          alternative = "twosided", method = "theoretical")
```


H0: µ{mu_nonsmoker} - µ{mu_smoker} = 0, There is no difference in the mean of the length of pregnancy in weeks between a smoker and nonsmoker mothers

HA: µ{mu_nonsmoker} - µ{mu_smoker} != 0, There is a difference in the mean of the length of pregnancy in weeks between a smoker and nonsmoker mothers



Calculate a 95% confidence interval how length of pregnency in weeks depend on the status of the mother as a smoker or nonsmoker.

```{r}

inference(y = nc$weeks, x = nc$habit, est = "mean", type = "ci", null = 0, 
          alternative = "twosided", method = "theoretical")

```

Conclusion

Given a p-value of  0.6038 , we affirm the null hypothesis and fail to reject the null hypothesis determine that there is no significant difference between mean of length of pregnancy in weeks between a smoker and nonsmoker mothers.


